%% acknowledgments.tex

% From mitthesis package
% Version: 1.01, 2023/10/16
% Documentation: https://ctan.org/pkg/mitthesis


\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

To begin, I would like to acknowledge that the analysis described in this thesis was a large effort carried out by many others at CERN beyond just myself. My primary focus during the analysis was in developing the muon fake rate, helping build the multivariate analysis tools, and analyzing backgrounds, which wouldn't have culminated to much had it not been for the work of others. To that end, I'd like to specifically thank Dima Kovalskyi, Christoph Paus, Zhangqier Wang, Kai-Feng Chen, Wei-Hsiang Huang, and Priyanka Sadangi for their contributions to the rest of the analysis. I hope this thesis properly represents the results of all your hard work. 

\vspace{5mm}

I would like to specifically thank Christoph for his amazing role as a mentor to me during my time here at MIT. He certainly has had an enormous impact in kindling my curiosity and fascination of physics. 

\vspace{5mm}

I'd also like to specifically thank Zhangqier for the many, many hours he has invested in me and for all his patience as I face the struggles an undergraduate typically faces while trying to understand particle physics. 

\vspace{5mm}

The friends I've made at the PPC group have made every day I spent in lab enjoyable and memorable. Thanks Luca, Pietro, Ludo, Marti, Kyungseop, and Chad for all the good times.

\vspace{5mm}

I thank all the other groups I've briefly been a part of, including ETP at KIT and Phillip Harris' group at MIT. I've profited significantly from the diversity of talented mentorship I've received over the years from all of you.

%The old groups I've been a part of: MIT (Dylan Rankin and Phillip Harris) and ETP (Markus Klute, Minseok, Jan, Joost, Neils, etc.)

\vspace{5mm}

My profound appreciation goes to the many professors and TAs who spent countless hours trying to teach me physics. A special thank you to Daniel Harlow, Netta Engelhardt, and Sam Alipour-fard for their amazing commitment to teaching and unbelievable insights. 

\vspace{5mm}

Of course, no acknowledgment would be complete without thanking my family Vivian, Constanze, and Markus for their endless love and support; I am nothing if not for y'all. 

\vspace{5mm}

Lastly, my love for physics has come from a love of learning from and hanging out with the people I have met along the way. Though I cannot name everyone here, I am deeply grateful to all of you.
